---
title: "Ridge Regression Overview"
author: "Naomi Edegbe"
date: "2024-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(car) #VIF function
library(MASS) #Ridge regression
```


Reference for this analysis is provided in Ch. 9 Special Topics in Regression, Section 9.7 Ridge Regression page 508 in Textbook.

## Data
The data in the FTCCIGAR file contains the tar content (x1), nicotine content (x2), weight (x3), and carbon monoxide content (y) for each in a sample of 25 cigarette brands.
```{r}
cigar <- read_csv("FTCCIGAR.csv")
head(cigar)
```

## Model
E(y) = β0 + β1x1 + β2x2 + β3x3 and discovered several signs of multicollinearity.
```{r ols model}
lm1 <- lm(co ~ tar + nicotine + weight, data = cigar)
summary(lm1)
```
The model coefficients match the SAS printout :)

Next, to check the VIF
```{r vif}
vif(lm1)
```
As the VIF of tar and nicotine are high (>10), we can conclude their is an issue of multicollinearlity present in our data.

```{r ridge model}
lmr <- lm.ridge(co ~ tar + nicotine + weight, data = cigar, lambda = 0.1) #c = 0.1 in text
lmr
```

```{r}
lmr3 <- lm.ridge(co ~ tar + nicotine + weight, data = cigar, lambda = seq(0,10,1)) 
lmr3$GCV
```

```{r method in meeting}
lambda1 <- seq(0,10,0.01)
lmr4 <- lm.ridge(co ~ tar + nicotine + weight, data = cigar, lambda = lambda1)
gcv <- lmr4$GCV

i = 1:length(gcv)
location = i[gcv[i]==min(gcv)]
opt.lambda = lambda1[location]
opt.lambda
```

```{r}
lmr5 <- lm.ridge(co ~ tar + nicotine + weight, data = cigar, lambda = opt.lambda)
lmr5
```




Plot ridge regression with multiple c values
```{r}
plot(lm.ridge(co ~ ., cigar,
              lambda = seq(0,0.1,0.001)))
```

```{r}
select(lm.ridge(co ~ ., cigar,
              lambda = seq(0,0.1,0.001)))
```

# Modeling with lmridge package
https://journal.r-project.org/articles/RJ-2018-060/RJ-2018-060.pdf

```{r}
library(lmridge)
```

```{r}
lmr2 <- lmridge(co ~., data = cigar, scaling = "sc", K = c(0,0.1,0.001))
print(lmr2) #unscaled 
```
The coefficients for K = 0.1 matches the SAS printout.
```{r}
summary(lmr2)
```


```{r}
vif(lmr2)
```
We can see at k = 0.1 that the vif estimates are close to 1. This results also matches the SAS printout. 

```{r}
plot(lmr2)
```
```{r correlation matrix}
# computing correlation matrix
cor_data = cor(cigar)
 
print("Correlation matrix")
print(cor_data)
```

# Modeling with glmnet package
